Using reference models in feature selection problems \vspace{.1in}

### 2019-10-19 paper/ref_approach.tex
rewrite Section 4 and 5

\subsection{Example of model selection with or without reference model}

@@

The remainder of the paper is structured as follows. In Section
\ref{reference-model-approach}, we review the concept of the reference
model, its benefits with examples and how it can be used as a filter
on data in a simple way. In Section \ref{comparison}, we show the
benefits of a reference model approach, regardless of what specific
procedure is applied, using the normal means estimation problem as a
benchmark. Finally, we end with a conclusion in Section
\ref{conclusion}.

@@

When using a reference model, we first obtain predictions for $f$
using all the covariates $\{X_{j}\}_{j=1}^{p}$ taking into account
that we have only observed the noisy representation $Y$ of $f$. If our
model is good, we are able to describe $f$ better than $Y$ itself can,
which improves the accuracy in the estimation of the relevance of the
covariates.  Figure \ref{fig:correlation} illustrates this process in
the form of a scatter plot of (absolute) correlations of the
covariates with $Y$ against the corresponding correlations with the
predictions of a reference model (in this case the posterior
predictive means of model \eqref{eq:ref_mod}; see Section
\ref{simulated-data}). Looking at the marginal distributions, we see
that using a reference model to filter out noise in the data, the two
groups of features (relevant and non-relevant) can be distinguished
much better than when the correlation is computed using the observed
noisy data directly.

@@

% \hypertarget{comparison}{%
%   \section{A comparison in the normal means problem framework}\label{comparison}}
\section{Complete feature selection}\label{comparison}

Alternative to minimal subset feature selection is \emph{complete
  feature selection}, in which the goal is to find all relevant
features which have some predictive information about the target. In
complete feature selection it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer
these effects.

The projection predictive approach was designed for the minimal subset
feature selection, but here examine its performance also in complete
feature selection. In addition, we analyse the benefits of reference
model approach in combination of a few common complete feature
selection approaches.

\subsection{Complete feature selection approaches to be compared}

Many Bayesian approaches for complete feature selection first fit a
Bayesian regression model using some sparsifying prior on the
regression coefficients, for example, as discussed in
\cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and
\cite{johnstone2004needles}.  In our experiments, we include the
Dirichlet-Laplace \cite[DL;][]{paper:dirichlet_laplace} prior and the
regularised horseshoe (RHS) prior \citep{paper:rhs}, which we think
scales better to higher number of covariates as compared to other
horseshoe-like priors.
\note{Aki: How this paragraph connects to the next paragraph?}

@@

Within the set of complete selection procedures, we consider three
Bayesian methods.  \cite{johnstone2004needles} use a spike-and-slab
type prior given by a mixture of a delta distribution in zero and a
heavy-tailed distribution \cite[see
also][]{paper:spike_slab_mitchell}. When the latter is a Laplace
distribution, they show an interesting thresholding property of the
median estimator and, based on this observation, present a complete
selection procedure using their prior. We refer to this method as
\emph{EB.med}. Further, we use a method to control for the local false
discovery rate (\emph{loc.fdr}) as described in \cite{paper:efron,
  efron2012large}. Finally, we device a simple selection procedure
based on posterior credible intervals using the regularised horseshoe
prior in which we select covariates whose 90\% credible intervals
excludes zero (\emph{ci.90}) \note{Reference for this?}.

All of these methods provide a complete selection procedure and we
compare their performance with and without using a reference model.
\note{Aki: Explain here how the reference mdoels are used for these. Point
  estimates replace data?}

As criteria of selection performance, we evaluate the average false
discovery rate (i.e., the ratio of the number of non-relevant selected
features over the number of selected features) and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \cite{paper:stability}. Table \ref{tab:comparison}
summarises the methods used in the numerical experiments.

\note{Aki: ``Type: Complete selection'' doesn't make sense in the table. Maybe we need to discuss when to use the term complete selection and what is the alternative title for this section.}

\begin{table}[tp]
\scriptsize
\centering
\begin{tabular}{l|l|l|l}
 Method & Abbreviation & Type & Quantities of interest \\ 
  \hline
 Regularised Horseshoe & RHS & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Dirichlet-Laplace & DL & Shrinkage prior & SSE, SESE, $\kappa$ \\
 Local false discovery rate & loc.fdr & Complete selection & fdr, sensitivity, stability \\
 Median estimator & EB.med & Complete selection & fdr, sensitivity, stability \\
 Credible intervals & ci.90 & Complete selection & fdr, sensitivity, stability \\
\end{tabular}
\caption{Summary of the shrinkage and selection methods used in the numerical experiments.}
\label{tab:comparison}
\end{table}

\hypertarget{complete-selection}{%
\subsection{Details of methods?}\label{complete-selection}}

\note{Should this be combined with previous subsection?}

As complete selection procedures we consider the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, the selection
by posterior credible intervals and an adaptation of the projection 
predictive approach that we name iterative projection.\

The control of the local false discovery rate is applied through the
\texttt{R}-package \texttt{locfdr}. It consists of testing the
z-values $\{z_{j}\}_{j=1}^{p}$ on whether they belong to the
theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$
meaning no relevance) against the alternative hypothesis distribution
$f_{1}$. In our case $f_{0}$ corresponds to the standard normal
distribution (see expression \eqref{eq:normal_means_problem2}). The
quantity of interest is the local false discovery rate (loc.fdr)
defined as: \
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)},
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and
$f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal
distribution of the z-values. The latter is estimated using splines
with 7 degrees of freedom. We select features with local false
discovery rate below $0.2$, which is suggested by
\citet{efron2012large} as it corresponds to a Bayes factor larger than
36 (assuming $\pi_{0}\geq0.9$). The results of the comparison are not
sensitive to the specific value. We use the default setting to
estimate $\pi_{0}$ from the data provided by the package.

The empirical Bayes median procedure is given by the
\texttt{R}-package \texttt{EbayesThresh} and consists of fitting a
Bayesian model with a prior composed by a mixture of a delta in zero
and a heavy-tailed distribution. As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.

The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles. \note{Reference?}

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,..,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \}
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule.
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.
\hl{NOTE: if necessary I can add a pseudo-code illustration of the iterative projection algorithm, as 
I did in the thesis}

\subsection{Simulated data}

\note{Aki: Why two subsections on simulated data? Combine this with next one, or did I understand their contents wrong?}

The normal means problem consists of estimating the (usually
sparse) vector of means of a vector of normally distributed
observations. The dimensionality of the vector of means is denoted by
$p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the
random variables $\{Z_{j}\}_{j=1}^{p}$. The task is to estimate the
latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}), \quad j=1,..,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data, $Z$, in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples in \cite{paper:efron, efron2012large}.

In our experiments, we retrieve the normal means problem from the
sample correlations between the target and the covariates using the
Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a
continuous target random variable $Y$, a set of $p$ continuous
covariates $\{X_{j}\}_{j=1}^{p}$, and denote
$\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$
statistical units and define $r_{j}$ the sample correlation between
the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the
$j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the
Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as
$T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate
normally distributed, the corresponding transformed correlations are
approximately normally distributed with known variance: \
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p.
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and
denoting the results as $z_{j}$, we have the formulation
\eqref{eq:normal_means_problem} of the normal means problem, this time
with unit variance: \
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p.
\end{equation}
Now the quantities of interest $\theta_{j}$ are equal to
$\sqrt{n-3}\,T_{F}(\rho_{j})$.

We use different levels of correlation $\rho\in\{0.3,0.5\}$ and number
observation $n\in\{50,70,100\}$. The total number of covariates $p$
and the number of relevant covariates $k$ are fixed to $p = 1000$ and
$k = 100$, respectively. In general, the lower $\rho$ and $n$, the
more challenging the selection.

\hypertarget{simulated-data}{%
\subsection{Simulated data}\label{simulated-data}}

Here we use simulated data in order to control the difficulty of the
selection problem. We rely on the scheme \eqref{eq:simulated_data}
using different levels of correlation $\rho\in\{0.3,0.5\}$ and number
observation $n\in\{50,70,100\}$. The total number of covariates $p$
and the number of relevant covariates $k$ are fixed to $p = 1000$ and
$k = 100$, respectively. In general, the lower $\rho$ and $n$, the
more challenging the selection. As mentioned, this data generation
mechanism was originally proposed by \cite{paper:projpred} who also
devise a reference model performing sufficiently well in this
situation. It consists of a linear regression using the first five
supervised principal components (SPCs) \citep{bair2006prediction,
  piironen2018} as covariates and imposing an hierarchical prior on
their coefficients: \
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10). \\
\end{aligned}
\end{equation}
In the above, $u_{ij}$ represents the $j$-th SPC evaluated at
observation $i$, and $s_{max}$ denotes the sample standard deviation
of the largest SPC.  The SPCs are computed through the
\texttt{R}-package \texttt{dimreduce} setting the screening threshold
parameter at $0.6s_{max}$.  In our experiments the results are not
sensible to such a choice of the screening threshold, yet a more
principled approach would be to use cross-validation to select the
threshold as done in \cite{paper:projpred}.

The normal means problem formulation is obtained as explained in
Section \ref{comparison} by applying the Fisher-z-transformation to
the empirical correlations. We label the reference approach as
``ref'', which consists of using the reference model on top of the
computation of the z-values, and the standard approach based on the
observed data simply as ``data''. In Section \ref{shrinkage-signal},
we analyse the different shrinkage and signal estimates in the context
of the full Bayes estimation methods by comparing the ref and data
approach. In Section \ref{complete-selection}, we present results for
the complete selection procedures.

@@

\subsection{Simulation results of complete feature selection}

@@

\hypertarget{real-world-data}{%
\subsection{The body fat data: A real world example}\label{real-world-data}}

\note{Aki: Is this only complete feature selection or also more results for the minimal subset feature selection?}

We conclude our  experiments using the real world `body fat' dataset
one more time, which was already described in the motivational example
comparing the projection predictive approach and the stepwise
regression in Section \ref{projection}. We carry out two different
analyses: one using the normal means problem framework and another one
using selection via stepwise backward regression. In the former, we
add noisy uncorrelated covariates to the original data up to a total
of 1000 features. These covariates are normally distributed and scaled
as the original covariates. We compute correlations between each
variable and the target variable, that is the amount of fat, and
transform them by Fisher-Z-transformation. The original assumption in
order for \eqref{eq:fisher_transformation} to hold is that the
variables are jointly normally distributed. In our experience the
normal approximation in \eqref{eq:fisher_transformation} is still
reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the
variance to be one, and instead estimate it from the data. The methods
we compare are those used in Section \ref{complete-selection}: the
control of the local false discovery rate (loc.fdr), the empirical
Bayes median (EB.med) and the selection by posterior credible
intervals at level 90\% (ci.90).  In order to vary the difficulty of
the selection, we bootstrap subsamples of different sizes, going from
$n=50$ up to $n=251$ (i.e., the full size of the data). For each
condition, results are averaged over 100 bootstrap samples of the
respective size.

@@

In the second analysis, we repeat the selection of Section
\ref{projection} via stepwise backward regression. In this case, the
overall number of covariates (original plus noisy) is 100, as it was
in the last part of Section \ref{projection}. We compare results with
and without using a reference model on top of the procedure following
our simple reference model approach outlined in
\eqref{eq:ref_mod}. Figure \ref{fig:bodyfat_step_refvsdata} shows the
number of irrelevant features included in the final model and the
out-of-sample root mean square error (RMSE). Results are based on 100
bootstrap samples on the whole dataset, and the predictive performance
is tested on the observations excluded at each bootstrap sample. We
observe that the reference model reduces the number of irrelevant
features included in the final model. This leads to less overfitting
and thus to improved out-of-sample predictive performance in terms of
RMSE. The reference model approach applied to the stepwise backward
regression achieves outstanding improvements considering its
simplicity, yet it does not reach the goodness of more thoughtful
implementations, for instance, the projective prediction approach (see
results of Section \ref{projection}).

@@

In this paper, we demonstrated the benefits of using a reference model
to improve feature selection, or more generally, model reduction. The
goal of this paper is not to provide a specific procedure to carry out
the selection, but rather to explain and motivate the general benefits
of a reference model.  We have seen how the reference model acts as an
approximation of the data generation mechanism through its predictive
distribution. Such approximation is generally less noisy than the
sample estimation available purely from the observed data, leading to
the main benefits of the reference model approach. In our comparisons,
we have analysed the effect of a reference model in the form of a
filter on the observed target values on top of different widely used
variable selection methods. Overall, using a reference model leads to
more accurate and stable selection results independently of the
specific selection method. These benefits apply to a large family of
different approaches all involving a reference model in one way or the
other. Some of these approaches have been present in the literature
for several years \cite[e.g. see][]{vehtari2012survey,paper:projpred}
but often without a clear explanation of \emph{why} they are actually
favourable and how they connect to other related approaches. We hope
that the present paper can fill some of these gaps by providing a
unifying framework and understanding of reference models.

We argue that, whenever it is possible to come up with a reasonable
reference model, it should be employed on top of the preferred
selection procedure or as an integral part of more complex methods,
for example, the projective prediction approach
\citep{paper:projpred}. Note that one of the main challenges in many
real world application will consist in devising a sensible reference
model itself and assessing its predictive performance. We did provide
a generic way to build simple reference models, but those may not
necessarily be appropriate in all situations. If more sophisticated
reference models are required, which are more specifically tuned to
the data and problem at hand, we recommend them to be developed using
a robust (Bayesian) modelling workflow, for instance, as outlined in
\cite{gabry2019visualization}.
\\

>>>>

\subsection{Example of model selection with or without reference model}
\label{intro-example}

@@

The remainder of the paper is structured as follows. In Section
\ref{reference-model-approach}, we review the concept of the reference
model, its benefits with examples and how it can be used as a filter
on data in a simple way. In Section \ref{comparison-minimal-subset} and 
Section \ref{comparison-complete-subset}, we show the
benefits of a reference model approaches for minimal and complete 
feature selection, respectively, before we illustrate the use 
of a reference model in a detailed case study of the 'body fat' data.
Finally, we end with a conclusion in Section \ref{conclusion}.

@@

When using a reference model, we first obtain predictions for $f$
using all the covariates $\{X_{j}\}_{j=1}^{p}$ taking into account
that we have only observed the noisy representation $Y$ of $f$. If our
model is good, we are able to describe $f$ better than $Y$ itself can,
which improves the accuracy in the estimation of the relevance of the
covariates.  Figure \ref{fig:correlation} illustrates this process in
the form of a scatter plot of (absolute) correlations of the
covariates with $Y$ against the corresponding correlations with the
predictions of a reference model (in this case the posterior
predictive means of model \eqref{eq:ref_mod}; see Section
\ref{simulations}). Looking at the marginal distributions, we see
that using a reference model to filter out noise in the data, the two
groups of features (relevant and non-relevant) can be distinguished
much better than when the correlation is computed using the observed
noisy data directly.

@@

% \hypertarget{comparison}{%
%   \section{A comparison in the normal means problem framework}\label{comparison}}
\section{Complete feature selection}\label{comparison-complete-subset}
\label{complete-selection}

An alternative to minimal subset feature selection is \emph{complete
  feature selection} in which the goal is to find all relevant
features that have some predictive information about the target. In
complete feature selection it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
feature selection but we will generalized it for the complete
feature selection case. In addition, we analyse the benefits of the reference
model approach in combination with three other complete feature
selection approaches.

As criteria of selection performance, we evaluate the average false
discovery rate (i.e., the ratio of the number of non-relevant selected
features over the number of selected features) and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \cite{paper:stability}.

%Table \ref{tab:comparison} summarises the methods used in the numerical experiments.

\subsection{Iterative projections}

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,..,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \}
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule.
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.
\hl{NOTE: if necessary I can add a pseudo-code illustration of the iterative projection algorithm, as 
I did in the thesis}
TODO: add the pseudo code here

\subsection{Alternative complete feature selection approaches}

% Many Bayesian approaches for complete feature selection first fit a
% Bayesian regression model using some sparsifying prior on the
% regression coefficients, for example, as discussed in
% \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and
% \cite{johnstone2004needles}.  In our experiments, we include the
% Dirichlet-Laplace \cite[DL;][]{paper:dirichlet_laplace} prior and the
% regularised horseshoe (RHS) prior \citep{paper:rhs}, which in our experience
% scales better to higher number of covariates as compared to other
% horseshoe-like priors.
% \note{Aki: How this paragraph connects to the next paragraph?}

@@

%Within the set of complete selection procedures, we consider three
%Bayesian methods.  \cite{johnstone2004needles} use a spike-and-slab
%type prior given by a mixture of a delta distribution in zero and a
%heavy-tailed distribution \cite[see
%also][]{paper:spike_slab_mitchell}. When the latter is a Laplace
%distribution, they show an interesting thresholding property of the
%median estimator and, based on this observation, present a complete
%selection procedure using their prior. We refer to this method as
%\emph{EB.med}. Further, we use a method to control for the local false
%discovery rate (\emph{loc.fdr}) as described in \cite{paper:efron,
%  efron2012large}. Finally, we device a simple selection procedure
%based on posterior credible intervals using the regularised horseshoe
%prior in which we select covariates whose 90\% credible intervals
%excludes zero (\emph{ci.90}) \note{Reference for this?}.

We consider three alternative complete selection procedures,
 the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, and the selection
by posterior credible intervals.

The control of the local false discovery rate consists of testing the
z-values $\{z_{j}\}_{j=1}^{p}$ on whether they belong to the
theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$
meaning no relevance) against the alternative hypothesis distribution
$f_{1}$. In our case $f_{0}$ corresponds to the standard normal
distribution (see expression \eqref{eq:normal_means_problem2}). The
quantity of interest is the local false discovery rate (loc.fdr)
defined as: \
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)},
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and
$f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal
distribution of the z-values. The latter is estimated using splines
with 7 degrees of freedom. We select features with local false
discovery rate below $0.2$, which is suggested by
\citet{efron2012large} as it corresponds to a Bayes factor larger than
36 (assuming $\pi_{0}\geq0.9$). The results of the comparison are not
sensitive to the specific value. We use the default setting to
estimate $\pi_{0}$ from the data provided by the \texttt{R}-package
\texttt{locfdr} (cite), which implements the method.

The empirical Bayes median procedure consists of fitting a
Bayesian model with a prior composed by a mixture of a delta in zero
and a heavy-tailed distribution. It is implemented in the
\texttt{R}-package \texttt{EbayesThresh} (cite). As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.

The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles (cite).

All of these methods provide a complete selection procedure and we
compare their performance with and without using a reference model.
That is, in the data condition, we apply the method on the original
data $y$ while, in the referece model condition, we replace $y$ by
their mean predictions $\hat{y}$ based on the reference model.

%\note{Aki: ``Type: Complete selection'' doesn't make sense in the table. Maybe we need to discuss when to use the term complete selection and what is the alternative title for this section.}

% \begin{table}[tp]
% \scriptsize
% \centering
% \begin{tabular}{l|l|l|l}
% Method & Abbreviation & Type & Quantities of interest \\ 
%  \hline
% Regularised Horseshoe & RHS & Shrinkage prior & SSE, SESE, $\kappa$ \\
% Dirichlet-Laplace & DL & Shrinkage prior & SSE, SESE, $\kappa$ \\
% Local false discovery rate & loc.fdr & Complete selection & fdr, sensitivity, stability \\
% Median estimator & EB.med & Complete selection & fdr, sensitivity, stability \\
% Credible intervals & ci.90 & Complete selection & fdr, sensitivity, stability \\
% \end{tabular}
% \caption{Summary of the shrinkage and selection methods used in the numerical experiments.}
% \label{tab:comparison}
% \end{table}

\subsection{Simulations}
\label{simulations}

To investigate the performance of the four complete selection procedures,
we are going to use simulations based on the normal means problem.
The normal means problem consists of estimating the (usually
sparse) vector of means of a vector of normally distributed
observations. The dimensionality of the vector of means is denoted by
$p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the
random variables $\{Z_{j}\}_{j=1}^{p}$. The task is to estimate the
latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}), \quad j=1,..,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data, $Z$, in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples in \cite{paper:efron, efron2012large}.

In our experiments, we retrieve the normal means problem from the
sample correlations between the target and the covariates using the
Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a
continuous target random variable $Y$, a set of $p$ continuous
covariates $\{X_{j}\}_{j=1}^{p}$, and denote
$\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$
statistical units and define $r_{j}$ the sample correlation between
the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the
$j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the
Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as
$T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate
normally distributed, the corresponding transformed correlations are
approximately normally distributed with known variance: \
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p.
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and
denoting the results as $z_{j}$, we have the formulation
\eqref{eq:normal_means_problem} of the normal means problem, this time
with unit variance: \
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p.
\end{equation}
In our situation, the quantities of interest $\theta_{j}$ are equal to
$\sqrt{n-3}\,T_{F}(\rho_{j})$.

We use different levels of correlation $\rho\in\{0.3,0.5\}$ and number
observation $n\in\{50,70,100\}$. The total number of covariates $p$
and the number of relevant covariates $k$ are fixed to $p = 1000$ and
$k = 100$, respectively. In general, the lower $\rho$ and $n$, the
more challenging the selection. As mentioned, this data generation
mechanism was originally proposed by \cite{paper:projpred} who also
devise a reference model performing sufficiently well in this
situation. It consists of a linear regression using the first five
supervised principal components (SPCs) \citep{bair2006prediction,
  piironen2018} as covariates and imposing an hierarchical prior on
their coefficients: \
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10). \\
\end{aligned}
\end{equation}
In the above, $u_{ij}$ represents the $j$-th SPC evaluated at
observation $i$, and $s_{max}$ denotes the sample standard deviation
of the largest SPC.  The SPCs are computed through the
\texttt{R}-package \texttt{dimreduce} setting the screening threshold
parameter at $0.6s_{max}$.  In our experiments the results are not
sensible to such a choice of the screening threshold, yet a more
principled approach would be to use cross-validation to select the
threshold as done in \cite{paper:projpred}.

% The normal means problem formulation is obtained as explained in
% Section \ref{comparison} by applying the Fisher-z-transformation to
% the empirical correlations. We label the reference approach as
% ``ref'', which consists of using the reference model on top of the
% computation of the z-values, and the standard approach based on the
% observed data simply as ``data''. In Section \ref{shrinkage-signal},
% we analyse the different shrinkage and signal estimates in the context
% of the full Bayes estimation methods by comparing the ref and data
% approach. In Section \ref{complete-selection}, we present results for
% the complete selection procedures.

@@

@@

\hypertarget{real-world-data}{%
\section{Case study: Body fat data}\label{real-world-data}}

We conclude our  experiments using the real world `body fat' dataset
one more time, which was already described in the motivational example
comparing the projection predictive approach and the stepwise
regression in Section \ref{intro-example}. We carry out two different
analyses: one using the normal means problem framework and another one
using selection via stepwise backward regression. In the former, we
add noisy uncorrelated covariates to the original data up to a total
of 1000 features. These covariates are normally distributed and scaled
as the original covariates. We compute correlations between each
variable and the target variable, that is the amount of fat, and
transform them by Fisher-Z-transformation. The original assumption in
order for \eqref{eq:fisher_transformation} to hold is that the
variables are jointly normally distributed. In our experience the
normal approximation in \eqref{eq:fisher_transformation} is still
reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the
variance to be one, and instead estimate it from the data. The methods
we compare are those used in Section \ref{complete-selection}: the
control of the local false discovery rate (loc.fdr), the empirical
Bayes median (EB.med) and the selection by posterior credible
intervals at level 90\% (ci.90).  In order to vary the difficulty of
the selection, we bootstrap subsamples of different sizes, going from
$n=50$ up to $n=251$ (i.e., the full size of the data). For each
condition, results are averaged over 100 bootstrap samples of the
respective size.

@@

In the second analysis, we repeat the selection of Section
\ref{intro-example} via stepwise backward regression. In this case, the
overall number of covariates (original plus noisy) is 100, as it was
in the last part of Section \ref{intro-example}. We compare results with
and without using a reference model on top of the procedure following
our simple reference model approach outlined in
\eqref{eq:ref_mod}. Figure \ref{fig:bodyfat_step_refvsdata} shows the
number of irrelevant features included in the final model and the
out-of-sample root mean square error (RMSE). Results are based on 100
bootstrap samples on the whole dataset, and the predictive performance
is tested on the observations excluded at each bootstrap sample. We
observe that the reference model reduces the number of irrelevant
features included in the final model. This leads to less overfitting
and thus to improved out-of-sample predictive performance in terms of
RMSE. The reference model approach applied to the stepwise backward
regression achieves outstanding improvements considering its
simplicity, yet it does not reach the goodness of the much
more sophisticated projective prediction approach (see
results of Section \ref{intro-example}).

@@

In this paper, we demonstrated the benefits of using a reference model
to improve feature selection, or more generally, model reduction. 
We have motivated and explained the general benefits
of a reference model regardless of the method it is applied in combination
with. Specifically, we have seen how the reference model acts as an
approximation of the data generation mechanism through its predictive
distribution. Such approximation is generally less noisy than the
sample estimation available purely from the observed data, leading to
the main benefits of the reference model approach. In our comparisons,
we have analysed the effect of a reference model in the form of a
filter on the observed target values on top of different widely used
variable selection methods. Overall, using a reference model leads to
more accurate and stable selection results independently of the
specific selection method. These benefits apply to a large family of
different approaches all involving a reference model in one way or the
other. Some of these approaches have been present in the literature
for several years \cite[e.g., see][]{vehtari2012survey,paper:projpred}
but often without a clear explanation of \emph{why} they are actually
favourable and how they connect to other related approaches. We hope
that the present paper can fill some of these gaps by providing a
unifying framework and understanding of reference models.

We argue that, whenever it is possible to come up with a reasonable
reference model, it should be employed on top of the preferred
selection procedure or as an integral part of more complex methods,
for example, the projective prediction approach
\citep{paper:projpred}. Note that one of the main challenges in many
real world application will consist in devising a sensible reference
model itself and assessing its predictive performance. We did provide
a generic way to build simple reference models, but those may not
necessarily be appropriate in all situations. If more sophisticated
reference models are required, which are more specifically tuned to
the data and problem at hand, we recommend them to be developed using
a robust (Bayesian) modelling workflow, for instance, as outlined in
\cite{gabry2019visualization}.
